---
title: "Project 4 -- SQL"
description: Analysis of the data in the Wideband Acoustic Immittance (WAI) Database 
author: Natalie Sullivan Wu
date: November 26, 2024
format: html

execute: 
  warning: false 
  message: false
---

I will use the data from Wideband Acoustic Immittance (WAI) Database hosted by Smith College. I am using the data to recreate the figure "Mean Absorbance from Each Publication in the WAI Database" from Susan E. Voss Ph.D. It is originally from Resource Review. Ear and Hearing 40(6):p 1481, November/December 2019 (link: <https://journals.lww.com/ear-hearing/citation/2019/11000/resource_review.21.aspx>)

```{r}
library(RMariaDB)
library(tidyverse)
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname= "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)
```

I started this project by using the SQL query, where I grouped together the data by different categories of "Frequency", "Year", "Instrument", "Identifier", and "AuthorShortList". I joined together the PI info and Measurements table to get the AuthorShortList. The table demonstrated the mean absorbance of frequencies starting at 200 Hertz and ending at 8000 Hertz.

```{sql}
#| connection: con_wai
#| output.var: "new_table"
    SELECT 
        m.Frequency,
        m.Identifier,
        m.Instrument,
        AVG(m.Absorbance) AS Mean_Absorption, 
        CONCAT(pi.AuthorsShortList, ' (', pi.Year, ')', ' N=', COUNT(DISTINCT CONCAT(m.SubjectNumber, '-', m.Ear)),'; ', m.Instrument) AS Label
    FROM 
        Measurements m
    JOIN 
        PI_Info pi
        ON m.Identifier = pi.Identifier
    WHERE 
        m.Identifier IN ('Abur_2014', 'Feeney_2017', 'Groon_2015', 'Lewis_2015', 'Lui_2008', 'Rosowski_2012', 'Shahnaz_2006', 'Shaver_2013', 'Sun_2016', 'Voss_1994', 'Voss_2010', 'Werner_2010')
        AND m.frequency BETWEEN 200 AND 8000
    GROUP BY 
        m.Frequency, 
        m.Identifier, 
        m.Instrument
    ORDER BY 
        m.Frequency, 
        m.Identifier, 
        m.Instrument
```

There is a pattern where absorbance increases when frequency increases. However, after it reaches 1000 Hertz, it started to decrease and drop.

```{r}
library(ggplot2)
ggplot(new_table, aes(x = Frequency, y = Mean_Absorption, color = Label)) +
  geom_line(size = 0.6, na.rm = TRUE) +  
  scale_x_continuous(
    name = "Frequency (Hz)",
    breaks = c(0, 2000, 3000, 3500, 3850, 6000, 7000, 7500, 8000),
    labels = c("200", "400", "600", "800", "1000", "2000", "4000", "6000", "8000")
  ) +
  scale_y_continuous(
    name = "Mean Absorbance",
    limits = c(0, 1), 
    breaks = seq(0, 1, by = 0.1)  
  ) +
  labs(
    title = "Mean Absorbance From Each Publication in the WAI Database",  
    color = NULL 
  ) +
  theme_minimal() + 
  theme(
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 7),
    plot.title = element_text(size = 10, face = "bold", hjust = .5),
    legend.text = element_text(size = 5),
    legend.position = c(0.27, 0.9),
    legend.key.size = unit(0.1, "cm"),
    legend.spacing = unit(0.01, "cm"),
    legend.background = element_rect(color = "black", size = .2, fill = "white"),
    aspect.ratio = 1
  ) 
```

Next, I will do the same thing as before, however, will group by "sex" as well. I used the study Abur (2014) to measure and show the average absorbamce of frequencies.

```{sql}
#| connection: con_wai
#| output.var: "new_table4"
SELECT 
    m.Frequency, 
    s.Sex, 
    AVG(m.Absorbance) AS Avg_Absorbance
FROM 
    Measurements m
JOIN 
    Subjects s
ON 
    m.SubjectNumber = s.SubjectNumber
WHERE 
    m.Identifier = 'Abur_2014'
    AND s.Sex != 'Unknown'
    AND m.frequency BETWEEN 200 AND 8000
GROUP BY 
    m.Frequency, s.Sex
ORDER BY 
    m.Frequency, s.Sex;

```

The two different lines show two sexes as subjects. The plot shows how male and female subjects have similar absorbance rates.

```{r}
library(ggplot2)

ggplot(new_table4, aes(x = Frequency, y = Avg_Absorbance, color = Sex)) +
  geom_line(size = 0.7, na.rm = TRUE) +
  scale_x_log10(
    name = "Frequency (Hz)",
    breaks = c(200, 400, 600, 800, 1000, 2000, 4000, 6000, 8000),
    labels = c("200", "400", "600", "800", "1000", "2000", "4000", "6000", "8000")
  ) +
  scale_y_continuous(
    name = "Average Absorbance",
    limits = c(0, 1), 
    breaks = seq(0, 1, by = 0.1)
  ) +
  labs(
    title = "Average Absorbance by Frequency and Gender",
    color = NULL 
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 8),
    legend.position = c(0.1, 0.9),
    legend.box.background = element_rect(color = "black", fill = "white", size = 0.5) 
  ) 
```
