---
title: "Birthday Paradox Presentation"

author: "Natalie Sullivan Wu"
subtitle: "December 9, 2024"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
    code-block-height: 100px

execute:
  echo: true
  warning: false
  message: false
---
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
```

## Permutation Test

Birthday paradox -- Harold Davenport (1927)
- The birthday paradox (birthday problem) is the fact that there is a 50% or larger chance that two people in a group of 23 have the same birthday

# Steps 

## My Project
- find the birthday paradox for N amount of people
```{r}
sim_shared_bday <- function(N){
  bdays <- sample(1:365, N, replace = TRUE)
  return(length(bdays) != length(unique(bdays)))
}
```

## Estimating the Probability of Shared Birthdays
- multiple simulations using sapply()
```{r}
estimate_shared_bday <- function(N, num_sims){
  outcome <- sapply(1:num_sims, function(x) sim_shared_bday(N))
  return(mean(outcome))
}
```

## Set the Parameters, Define our Probability, and Plot

```{r}
grp_size <- 2:100
num_sims <- 1000 
prob <- map_dbl(grp_size, ~ estimate_shared_bday(.x, num_sims))
data <- data.frame(group_size = grp_size, probability = prob)

ggplot(data, aes(x = group_size, y = probability)) + 
  geom_point(color = "red", size = 2) + 
  geom_line(size = 1) + 
  labs(
    title = "Shared Birthday Probaility Graph",
    x = "Amount of People",
    y = "Probability of Having a Shared Birthday"
  ) + 
  theme_minimal()
```

## In Comparison to the 50% Threshhold

```{r}
grp_size <- 2:100
num_sims <- 1000 
prob <- map_dbl(grp_size, ~ estimate_shared_bday(.x, num_sims))
data <- data.frame(group_size = grp_size, probability = prob)

ggplot(data, aes(x = group_size, y = probability)) + 
  geom_point(color = "red", size = 2) + 
  geom_line(size = 1) + 
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "blue", size = 1) + 
  scale_x_continuous(breaks = seq(0, 100, by = 5)) + 
  labs(
    title = "Shared Birthday Probaility Graph",
    x = "Amount of People",
    y = "Probability of Having a Shared Birthday"
  ) + 
  theme_minimal()
```

## Takeaways 
- Increase in group size increases --> rapid increase in probability (exponential) 
- Contradicts intuitive prediction --> The probability would increase slowly 
- After exponential growth --> Slope decreases (at 50 people)
