---
title: "Rodent TidyTuesday Analysis"

description:
  Analysis of Species of Rodents data from TidyTuesday
author: Natalie Sullivan Wu
date: September 18, 2024
format: html

execute: 
  warning: false 
  message: false
---

```{r}
rodentdata <- tidytuesdayR:: tt_load('2023-05-02')
species <- rodentdata$species
```

This data is from the Portal Project (link: <https://portal.weecology.org/>). The [Weecology research group](https://www.weecology.org/) is a long term ecological research site that monitors rodents, plants, ants, and weather. This dataset serves as the core dataset for the [Data Carpentry Ecology](https://datacarpentry.org/ecology-workshop/) material.

```{r}
library(dplyr)
weight <- species %>%
  mutate(difference_weight = meanwgt - juvwgt, )
```

```{r}
library(ggplot2)
ggplot(weight, aes(x = reorder(commonname, difference_weight), y = difference_weight, fill = difference_weight))+ 
  geom_bar(stat = "identity")+
  coord_flip() +
  scale_fill_gradient(low = "yellow", high = "darkblue") + 
  labs(title = "Difference in Weight for Rodent Species", 
       x = "Name of Species",
       y = "Difference in Weight"
       ) + 
  theme_minimal()
```

The white-throated woodrat had the highest difference in weight, at nearly 80, out of all the species of rodents. This is interesting, considereing every other species of rodent has a signicantly smaller weight
